---
output: github_document
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/muschellij2/SummarizedActigraphy/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/muschellij2/SummarizedActigraphy/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/muschellij2/SummarizedActigraphy/branch/master/graph/badge.svg)](https://codecov.io/gh/muschellij2/SummarizedActigraphy?branch=master)
[![Codecov test coverage](https://codecov.io/gh/muschellij2/SummarizedActigraphy/graph/badge.svg)](https://app.codecov.io/gh/muschellij2/SummarizedActigraphy)
<!-- badges: end -->

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  cache = TRUE
)
```

# SummarizedActigraphy Package: 
The goal of `SummarizedActigraphy` is to provide functions for reading Actigraphy data and turn it into `SummarizedExperiment`s.

## Installation

You can install `SummarizedActigraphy` from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("remotes")
remotes::install_github("muschellij2/SummarizedActigraphy")
```

## Reading in some data

```{r}
library(SummarizedActigraphy)
url = paste0("https://github.com/THLfi/read.gt3x/files/",
             "3522749/GT3X%2B.01.day.gt3x.zip")
destfile = tempfile(fileext = ".zip")
dl = utils::download.file(url, destfile = destfile)
gt3x_file = utils::unzip(destfile, exdir = tempdir())
gt3x_file = gt3x_file[!grepl("__MACOSX", gt3x_file)]
path = gt3x_file
result = summarize_actigraphy(path)
result
```


## Reading in GT3X files

In GT3X files, you must be aware of [Idle Sleep Mode](https://actigraphcorp.my.site.com/support/s/article/Idle-Sleep-Mode-Explained), which saves power on the device, but the device essentially stops recording.  In most cases, you still want a "full" time series without missing elements.
To read in `gt3x` files, we use `read.gt3x::read.gt3x`. By default `read.gt3x::read.gt3x`, that data isn’t in the output data set. If you use the `imputeZeros = TRUE` argument, then they are there, but all 0s, which doesn’t make sense (no gravity even!?). 

In Actigraph's ActiLife, the raw output data repeats the same value before the device went into sleep mode.  To mimic this behavior, the `SummarizedActigraphy::fix_zeros` function sets the values in the rows with all `0` to be `NA` (just for `X/Y/Z`, not `time`), and then uses `zoo::na.locf` for last observation carried forward (LOCF).  Aside: you could also use `tidyr::fill(direction = "down")` for the tidyverse folks.

In `SummarizedActigraphy::read_actigraphy`, we use `read.gt3x::read.gt3x(asDataFrame = TRUE, imputeZeroes = TRUE)` as the default.

See below as an example.

```{r}
data = SummarizedActigraphy::read_actigraphy(path)
df = data$data
df = tibble::as_tibble(df)
all_zero = df$X == 0 & df$Y == 0 & df$Z == 0
df[all_zero, ] %>% 
  dplyr::mutate(time = lubridate::floor_date(time, "1 sec")) %>% 
  dplyr::distinct()
df = SummarizedActigraphy::fix_zeros(df)
df[all_zero, ] %>% 
  dplyr::mutate(time = lubridate::floor_date(time, "1 sec")) %>% 
  dplyr::distinct()
```

Note well, the `fix_zeros` does not do anything with data that is all zero in the beginning of a time series (as there is no observation to carry forward).  This behavior should mimic ActiLife to our knowledge and that data is likely to be discarded regardless.

## Converting to wide 1440 format

```{r make1440}
hms_times = structure(seq(0, 86340, by = 60), class = c("hms", "difftime"),
                      units = "secs")
hms_times = tibble::tibble(time = hms_times)
measure = "AI_mean"
tmeasure = c("time", measure)
x = result[, tmeasure, drop = FALSE]
x = tibble::as_tibble(x)
x = dplyr::left_join(hms_times, x)
head(x)
x1440 = tidyr::spread(x, time, value = measure)
x1440
```
