---
title: "limmi"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{limmi}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(RNifti)
library(limmi)
```

# Data

The data is from https://www.fil.ion.ucl.ac.uk/spm/data/face_rfx/.  The data was from
the study of

> Henson, R. N. A., et al. "Face repetition effects in implicit and explicit memory tests as measured by fMRI." *Cerebral Cortex* 12.2 (2002): 178-186.

We have included a subset of this data in the `inst/extdata/` directory (from the `can` contrast):

```{r}
tarfile = system.file("extdata", "can.tar.gz", package = "limmi")
tarfile
exdir = tempfile()
dir.create(exdir)
files = untar(tarfile = tarfile, list = TRUE, exdir = exdir)
files = files[!grepl("^\\.", basename(files))]
unz = untar(tarfile = tarfile, files = files, exdir = exdir)
files = files[ grepl("hdr", basename(files))]
files
files = file.path(exdir, files)
```

# Create a Data Matrix

We will use the `nifti_images_to_matrix` function to read these images into a matrix:

```{r}
mat = nifti_images_to_matrix(files, verbose = FALSE)
attr(mat, "image_dim")
attr(mat, "image_pixdim")
dim(mat)
```

Let's read in one of the images and look at the contrast images:
```{r}
img = asNifti(files[[1]])
```
```{r}
have_neurobase = requireNamespace("neurobase", quietly = TRUE)
```
```{r}
pdim = c(3, pixdim(img), rep(1, 4))
if (have_neurobase) {
  neurobase::ortho2(img, pdim = pdim)
}
```

We can also look at the mean image over all voxels:
```{r}
mask = array(TRUE, dim = dim(img))
mean_image = limmi:::remake_nifti_image(rowMeans(mat), img, mask = mask >0)
if (have_neurobase) {
  neurobase::ortho2(mean_image, pdim = pdim)
}
```

Now, we could run `limma` directly on this matrix.  The `nifti_limma` function does this as a wrapper for the process.  We should likely have a mask file of the voxels we want to perform the operation, but otherwise it will work over all voxels, even those with `NaN` in all the regions. 

Let's make a mask of voxels where there is **any** non-missing in the data:

```{r}
keep = rowSums(!is.na(mat)) > 0
mask = array(keep, dim = dim(img))
mask = asNifti(mask, reference = img)
if (have_neurobase) {
  neurobase::ortho2(mask, pdim = pdim)
}
```

# Run `limma`

Now that we have the mask we want, let's run `limma` and a linear fit:
```{r}
fit = limma::lmFit(mat[keep, ])
```

## All in one line

We can use the `nifti_lmFit` function to perform this all in one function:
```{r}
check_fit = nifti_lmFit(imgs = files, mask = mask, verbose = FALSE)
all.equal(check_fit$coefficients, fit$coefficients)
```

# Running `eBayes`

Let's run `eBayes` with the defaults (not likely a good idea with this data), getting the results:
```{r}
eb = limma::eBayes(fit = fit)
coef_vec = eb$coefficients[,1]
coef_image = limmi:::remake_nifti_image(coef_vec, img, mask = mask > 0)
if (have_neurobase) {
  neurobase::double_ortho(mean_image, coef_image, pdim = pdim)
}
```


## All in one line
The `nifti_eBayes` function will do the reading, fitting, and `eBayes`:

```{r}
check_eb = nifti_eBayes(files, mask, verbose = FALSE)
all.equal(check_eb$empirical_bayes$coefficients, eb$coefficients)
```

```{r}
diff_from_mean = mean_image - coef_image
if (have_neurobase) {
  neurobase::ortho2(diff_from_mean, pdim = pdim)
}
hist(diff_from_mean)
```

The output of `nifti_eBayes` does the remaking of images, and the `images` element is a list of these output images from `eBayes`:

```{r}
if (have_neurobase) {
  neurobase::ortho2(check_eb$images$p.value, pdim = pdim)
}
```

By default, `nifti_eBayes` also does a `topTable` from `limma` to adjust the p-values, using a Benjamini-Hochberg adjustment:

```{r}
if (have_neurobase) {
  neurobase::ortho2(check_eb$images$adjusted_p_value, pdim = pdim)
  neurobase::ortho2(check_eb$images$adjusted_p_value <= 0.05, pdim = pdim)
}
```

Here we can compare the 2 maps, before and after adjustment:

```{r}
if (have_neurobase) {
  neurobase::ortho2(check_eb$images$p.value <= 0.05, pdim = pdim)
  neurobase::ortho2(check_eb$images$adjusted_p_value <= 0.05, pdim = pdim)
}
```
