---
title: "Summarizing Actigraphy Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Summarization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = ""
)
```

```{r setup}
library(SummarizedActigraphy)
```

# Data

The data is from https://github.com/THLfi/read.gt3x/files/3522749/GT3X%2B.01.day.gt3x.zip.  It is a daily GT3X file from an ActiGraph. 

Let's download the data:

```{r}
url = "https://github.com/THLfi/read.gt3x/files/3522749/GT3X%2B.01.day.gt3x.zip"
destfile = tempfile(fileext = ".zip")
dl = utils::download.file(url, destfile = destfile)
gt3x_file = utils::unzip(destfile, exdir = tempdir())
gt3x_file = gt3x_file[!grepl("__MACOSX", gt3x_file)]
path = gt3x_file
```

This data represents sub-second level accelerations in the X, Y, and Z directions.  Additional information from devices can be measured, such as temperature or light.  We will focus only on the accelerometry data.  The `GGIR::g.calibrate` function is a method to calibrate the ENMO values [@GGIR_calibrate].  Other types of activity data, would be things such as activity counts, step counts, or previously summarized data.  Data such as this is commonly calculated using proprietary methods or algorithms.  

# Create a Data Matrix

We will use the `read_actigraphy` function to read these files into an `AccData` object:

```{r}
x = read_actigraphy(path)
class(x)
names(x)
```

The `read_actigraphy` function uses the `read.gt3x::read.gt3x` for gt3x files, and uses functions from the `GGIR` package [@GGIR].  

The output has a data matrix in the `data.out`, which has `X`, `Y`, and `Z` columns, with an additional `time` column, which is a date/time column.  Additionally, the `header` object has additional metadata about the object:


```{r}
x$header
```

The sampling frequency is embedded in the header, but also found in the `freq` element:
```{r}
x$freq
```

In this case, there are `r x$freq` samples per second.

# Summarizing the data: Day-Second Level

The `summarize_daily_actigraphy` summarizes an `AccData` into second-level data for each day.   The output is an `tsibble`:

```{r}
daily = summarize_daily_actigraphy(x)
head(daily)
```

The process is as follows, we use `floor_date` from `lubridate`, so that each time is rounded to 1 seconds.  This rounding is what allows us to use `group_by` on the time variable (which is really date/time variable) and summarize the data:

```{r}
library(dplyr)
data = x$data.out
data = data %>% 
  mutate(time = lubridate::floor_date(time, unit = "1 seconds"),
         enmo = sqrt(X^2 + Y^2 + Z^2) - 1)
head(data)
```

In `summarize_daily_actigraphy`, the following process is done:

```{r}
data = data %>%
  group_by(time) %>%
  summarize(
    mad = (mad(X, na.rm = TRUE) + 
             mad(Y, na.rm = TRUE) +
             mad(Z, na.rm = TRUE))/3,
    ai = sqrt((var(X, na.rm = TRUE) + 
                 var(Y, na.rm = TRUE) + 
                 var(Z, na.rm = TRUE))/3),
    n_values = sum(!is.na(enmo)),
    enmo = mean(enmo, na.rm = TRUE)
  )
head(data)
```

This assumes that all non-NA values are valid, no wear-time estimation is done.  Also, estimating non-wear time/areas from this summarized data.  The `accelerometry::weartime` function estimates wear time, but is based on count values.  

This daily-level data is good for thresholding activity, such as into categories like vigorous activity.  Also, this data is used for estimating sedentary and active bouts, and transition probability between them.   

## ENMO 

One of the statistics calculated are ENMO, the Euclidean Norm Minus One, which is:

$$
ENMO_{t} = \frac{1}{n_{t}} \sum_{i=1}^{n_t} \left(\sqrt{X_{i}^2 + Y_{i}^2 + Z_{i}^2} - 1\right)
$$
where $t$ is the time point, rounded to one second.  In this case, $i$ represents each sample, so there are `r x$freq` ENMO measures per second, which means the number of samples for that time point, $n_{t}$, is `r x[["freq"]]`.  In most cases, however, ENMO and the other statstics are calculated at the 1-second level.  So the $ENMO_{t}$ from `summarize_daily_actigraphy` calculates $ENMO$ for each sample, then takes the mean $ENMO_{t}$ for 1-second intervals.  We will use $ENMO_{t}$ to represent this second-level data, leaving off any bar or hat, but in truth it is an average.  The subtraction of $1$ is to subtract $1$ gravity unit (g), from the data so it represents acceleration not due to gravity.


## Activity Index (AI)

The Activity Index (AI) was introduced by @bai.  It is calculated by:
$$
AI_{t} = \sqrt{\frac{Var(X_{t}) + Var(Y_{t}) + Var(Z_{t})}{3}}
$$

where 

$$
Var(X_{t}) = \frac{1}{n_{t}} \sum_{i=1}^{n_t} \left(X_{i} - \bar{X_{t}}\right)^2
$$

where $\bar{X_{t} = \frac{1}{n_{t}} \sum_{i=1}^{n_t} X_{i}$.  

## Median Absolute Deviation (MAD)

The Median Absolute Deviation (MAD) was introduced by XX.   It is calculated by:

$$
MAD_{t} = \frac{MAD(X_{t}) + MAD(Y_{t}) + MAD(Z_{t})}{3}
$$

where 

$$
MAD(X_{t}) = \text{median}_{t} \left|X_{i} - \text{median}(X_{t})\right|
$$
where $\text{median}_{t}$ is the median over all values $i = 1, \dots, n_{t}$, and $\text{median}(X_{t}) = \text{median}_{t} (X_{i})$. 


# Summarizing the data: Second Level

The `summarize_actigraphy` summarizes an `AccData` into second-level data for an "average" day or average profile.   For each second of the day, a summary statistic is taken, either the mean or the median.  The `summarize_actigraphy` gives both, and the output is an `tsibble`:

```{r}
average_day = summarize_actigraphy(x, .fns = list(mean = mean, median = median),)
head(average_day)
```

The number of rows is 86400, which is 60 seconds per minute, 60 minutes per hour, 24 hours per day: 

```{r}
nrow(average_day)
range(average_day$time)
```

These values represent an "average" day, where average is determined by mean and median.  

# Plotting an Average Day

Here we plot the AI values for each minute, summarized over mean and median:

```{r}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  library(ggplot2)
  library(magrittr)
  average_day %>%
    dplyr::rename_with(tolower) %>% 
    ggplot(aes(x = time, y = ai_mean)) +
    geom_line()
  
  average_day %>%
    dplyr::rename_with(tolower) %>% 
    ggplot(aes(x = time, y = ai_median)) +
    geom_line()
}
```

